Survival Analysis

### 2014-03-31 presentation.Rnw
Added descriptors slide
Added example graphs and confused myself
hazards is a rate, but of what to what?

@@

\begin{frame}{Data Collection Issues}
<<example-timeline, echo=FALSE, fig=TRUE>>=

@
\end{frame}

>>>>

<<echo=FALSE>>=
library(ggplot2)
library(survival)
library(KMsurv) # survival data sets
library(GGally) # survival curve support for ggplot2

data(bfeed)

# FDA recommends 6 months of breastfeeding ~ 27 weeks
bfeed.short <- subset(bfeed, duration < 32)
@

@@

\begin{frame}{Data Collection Issues}
\begin{center}
<<example-timeline, echo=FALSE, fig=TRUE, height=4, width=6>>=
timeline <- read.csv('example-timeline.csv', header=TRUE, row.names=1)

ggplot(timeline) +
  ggtitle('Example Event Timeline') +
  xlab('time') +
  ylab('observation') +
  labs(
    colour = "Observation",
    shape = "Event Occured?")+
  geom_segment(aes(
    colour = rownames(timeline),
    x = start,
    xend = end,
    y = rownames(timeline),
    yend = rownames(timeline))) +
  geom_point(aes(
    colour = rownames(timeline),
    shape = event,
    x = end,
    y = rownames(timeline)), size = 4) +
  geom_vline(
    xintercept = 6,
    colour = 'red',
    linetype = 'longdash')
@
\end{center}
\end{frame}

\begin{frame}{Descriptors}
Assume $f(t)$ is the instantaneous failure rate

\vspace{0.5in}

% Hazard stuff is wrong
\begin{tabular}{lllcc}
Survival
  & $P(\mbox{no event in } t_i < t)$
  & $S(t)$ & = & $\int^{t}_{0} f(x)$ \\
Hazard
  & $P(\mbox{no event until } t)$
  & $\lambda(t)$ & = & $^{f(t)} / _{S(t)}$ \\
Cum. \\ Hazard
  & E()
  & $\Lambda(t)$ & = & $\int^{t}_{0} \lambda(t) dt$ \\
\end{tabular}
\end{frame}

\begin{frame}{Descriptors}
\begin{tabular}{lcc}
  & independent & conditional \\
  instantaneous & $f(t)$ & $\lambda(t)$ \\
  cumulative & $S(t)$ & $\Lambda(t)$
\end{tabular}
\end{frame}

\begin{frame}{Example}
\begin{center}
<<bfeed-plain, echo=FALSE, fig=TRUE, height=4, width=6>>=
bfeed.fit <- survfit(Surv(duration, delta) ~ 1, data = bfeed.short)
ggsurv(bfeed.fit) +
  ggtitle('Breast Feeding Duration') +
  xlab('Duration in Weeks') +
  ylab('Survival') +
  geom_vline(aes(xintercept = 27),
    colour = 'red',
    linetype = 'longdash')
@
\end{center}
\end{frame}

\begin{frame}{Models}
(Cox) Proportional Hazards Model

$$\lambda(t | \mathbf{x}) = \lambda_{0}(t) \exp \left( \mathbf{x' \beta} \right)$$

Accelerated Failure Time Model

$$\log T = \mathbf{x' \beta} + \mathbf{\epsilon}$$

Exponential/Weibull Model

$$\lambda(t, \mathbf{x}) = \lambda_{0} \exp \left( \mathbf{x' \beta} \right)$$
\end{frame}
